Mining Scenario/Transaction Pairs in HW/SW Co-verification Using Symbolic Execution

### 2017-04-06 paper.tex
minor and comments

\end{enumerate}
%
%===============================================================================
\section{Working Example}
%{\bf Running example: }
%===============================================================================
In this section, we present an working example of FW and HW model to explain our 
co-verification methodologies.

@@

\textbf{Co-design Property} We verify that the transmitted data is the same as the 
received data in the {\em loopback} mode.  To do so, we annotate the FW model 
with an assertion $assert()$, as shown in Figure~\ref{firmware}.  Note that the 
array containers $rx\_b$ and $tx\_b$ stores the data obtained through the $inb$
and $outb$ function calls which directly read from and write to the UART IP ports
respectively. Section~\ref{property} describes the property specification in
HW-FW co-designs in more details. 
\[
  for(i=0; i<=N; i++) 
   \quad assert(rx\_b[i] == tx\_b[i]);
\]
%
\section{Scenario and Transaction}
%
We formally describe the notion of \emph{transaction} and \emph{scenario} with the
help of our working example. \\
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{|l|l|}
\hline
Wishbone Interface & Main Module \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
typedef unsigned char u8;
unsigned char inb 
 (unsigned long port) {
  return wb_read(port);
}
void outb (u8 value, 
 unsigned long port) {
  wb_write(port, value);
}
void wb_reset(void) {
 rst_i = 1;
 set_inputs();
 next_timeframe();
 rst_i = 0;
 stb_i = 0; 
 cyc_i = 0;
}
void wb_idle() {
 set_inputs();
 next_timeframe();
}
void wb_write(_u32 addr, 
     _u8 b) {
 adr_i = addr;
 dat_i = b;
 we_i = 1;
 cyc_i = 1;
 stb_i = 1;
 set_inputs();
 next_timeframe();
 we_i = 0;
 cyc_i = 0;
 stb_i = 0;
}
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
int main() {
wb_reset();
wb_idle();
// Configure the uart
outb (0x13, UART_MC);  
outb (0x80, UART_CM3);
outb (0x00, UART_CM2);
outb (0x00, UART_CM1);
outb (0x00, UART_CR);  
outb (0x03, UART_IE); 
char tx_b[] = "Hello world";
_u8 status = 0;
char rx_b[100];
int i=0,c=0,d=0;
// data transfer in loopback
for (i=0; i<1990; i++){
if (irq_o){
 status=inb(UART_IS)&0x0c;
 if(istatus==0x0c){
 //it was a tx_empty interrupt
 outb(*(tx_b+c),UART_TR); c++;
 }
 else{ //status==0x04
 //it was an rx_data interrupt
  rx_b[d] = inb(UART_TR); d++;
 }
}else {
 // no interrupt. 
 wb_idle();
 wb_idle();
 }
}
// property
for(i=0; i<=10; i++)
 assert(rx_b[i] == tx_b[i]);
}
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{Firmware model of UART IP}
\label{firmware} 
\end{figure}
%

\textbf{Scenario ($\mathcal{S}$):}
Let $B$ denote a set of all possible behaviors of a system. A scenario is a set of 
constraints $\mathcal{S}$ that restrict the system behaviors
to $B' \subseteq B$. For example, the left hand side of Figure~\ref{fig:st} 
gives an example scenario for the UART FW shown in Figure~\ref{firmware}. 
Here, $\mathcal{S}$ configure the UART IP in {\em loopback} mode with 
interrupt enabled.

@@

\textbf{Firmware Transaction ($\mathcal{FT}$):} 
A Firmware Transaction ($\mathcal{FT}$) is an untimed state machine 
with a start state and an end state, which is similar to the notion of 
transactions in TLM~\cite{codes14}. 
Each transition in $\mathcal{FT}$ is guarded with a quantifier free predicates
over FW variables and HW signals.
If the guard evaluates to $TRUE$, it may perform some action, which may be 
read/write data. If the guard is not specified, it is assumed to be $TRUE$.
$\mathcal{FT}$ is often manifested as control-flow paths in the firmware. 
The right hand side of Figure~\ref{fig:st} gives a 
$\mathcal{FT}$ for the UART firmware code shown in Figure~\ref{firmware}. 
It has a start state and an end state. $\mathcal{FT}$ is triggered from 
UART output signal, $irq_o$. An execution of $\mathcal{FT}$, also called a 
FW transaction instance, takes a simple path from start state to end state. 
The path taken by an execution is determined by the state of 
the FW variables or its interacting HW signals. Repeated execution of 
$\mathcal{FT}$ can happen only in a sequential manner.

@@

Let $\mathcal{G}$ be the set of quantifier free boolean predicates over 
state elements, $\{Reg\}$, $\{Mem\}$, $\{Wire\}$.  
A {\em transaction step}, $\varphi$, is defined by 2 tuple,
$\varphi \equiv \langle \Gamma_{\varphi}, \Delta_{\varphi} \rangle$, where 
$\Gamma_{\varphi} \equiv \{\alpha_1, \alpha_2, \ldots \alpha_{m}\}$
is a set of concurrent state updates to $\{Reg\}$, $\{Mem\}$, $\{Wire\}$ 
and $\Delta_{\varphi} \equiv \{\beta_1, \beta_2, \ldots \beta_{n}\} \subseteq
\mathcal{G}$.

A {\em Hardware Transaction} $(\mathcal{HT}) = (V, \Sigma, \rightarrow, init,
comp)$ is a labelled transition system, where $V$ is a set of transaction steps
$\varphi$ with a unique source vertex, $init \in V$, and a unique sink vertex, 
$comp \in V$, $\Sigma$ is a set of labels such that $\Sigma \subseteq 2^{G}$ and  
$\rightarrow \subseteq V \times \Sigma \times V$ is a set of labelled
transitions.

@@

\caption{Path-based HW/SW Co-verification Flow in \verifox

@@

\textbf{Step 3: Co-verification Model}
%
An unified co-verification model is constructed through appropriate composition of 
FW and its interacting HW threads.  This is done by instrumenting the 
software netlist model of HW RTL inside the FW thread at every invocation 
of the $next\_timeframe()$ primitive in interface functions.  Recall that a
call to $next\_timeframe()$ is used to advance the HW clock by one cycle.  A call
to the software netlist emulates the effect of one clock period in the RTL 
model. That is, the state of state holding elements (registers) and combinational 
elements (wires) are uniform in software netlist and RTL model after every invocation of 
$next_timeframe()$.  Furthermore, a software
netlist is a sequential model of HW RTL which is represented in C.  Thus, the
composition leads to a single-threaded sequential co-verification model in C.  
%
We now formally define a our co-verification model, denoted by 
$\mathcal{M}_{seq}$.
%
\[ 
\begin{array}[t]{@{}lll}
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{moduleName(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\ 
\end{array}
\]
%
Consider sets of expressions $Exp$ and Boolean expressions $BExp$ 
over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 
$(b \in BExp)$. A $\mathit{CStatement}$ is an assignment,
conditional, sequential concatenation or a loop.
$\mathit{moduleName(Var_1,\dots,Var_n)}$ invokes the top level 
procedure of $\mathcal{SN}$, which is also the top level 
module of the HW RTL, since the translation from RTL to software preserves
the module hierarchy.  The interface variables $\{Var\_1,\dots,Var\_n\}$ 
(input, output and inout ports) are declared inside the 
FW to communicate with $\mathcal{SN}$.  Figure~\ref{compose} shows the 
composition of $\mathcal{SN}$ and FW thread for the UART IP.
%
\begin{figure}[t]
\scalebox{.50}{\import{figures/}{compose.pspdftex}}
  \caption{Composition of FW and HW threads}
\label{compose}
\end{figure}

\textbf{Step 4: Property Driven Slicing of Co-verification Model}
A property-driven slicing is performed on the unified model, $\mathcal{M}_{seq}$.  
This step is purely syntactic, meaning that we perform a backward dependency 
analysis starting from the property which only preserve those 
program fragments that are relevant to a given property. The sliced program 
is then passed to the symbolic execution engine for path-based exploration. \\
%
%=======================================================================================
%\subsection{\verifox: Path-wise Symbolic Execution Tool} 
%=======================================================================================
\input{verifox}
%
\subsection{Approach 2: Monolithic Co-verification Technique}
%
\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.6]{figures/traditional_flow.pdf}
\caption{Monolithic HW/SW Co-verification Flow in \hwcbmcv}
\label{fig:conventional}
\end{center}
\end{figure}
%
\textbf{Step 1: Co-verification Model}
%
Figure~\ref{fig:conventional} illustrates the monolithic approach to HW/SW
co-veri\-fi\-cation, as implemented in \hwcbmcv.  
%\hwcbmcv is the only open-source
%formal co-verification tool that supports Verilog RTL (System Verilog 2009
%standard) and C/C++ firmware.  
In contrast to path-based approach,
\hwcbmcv maintains two separate flows for HW and FW.  The top flow in
Figure~\ref{fig:conventional} uses synthesis to obtain either a bit-level or
a word-level netlist from Verilog RTL.  A bit-level netlist is represented 
in And Inverter Graph (AIG) format.  Whereas, a word-level netlist is represented 
in format that resembles SMT-LIB standards.  The bottom flow illustrates the
translation of C program into static single assignment (SSA) form. 
Synthesizing netlist from RTL is a standard technique which is widely 
used in hardware model checkers~\cite{DBLP:journals/tse/Clarke76}.
Similarly, translating a C program into a SSA representation is a 
standard technique which is widely used in 
software verification tools, such as CBMC~\cite{cbmc.tacas:2004}. 
\hwcbmcv generates a monolithic co-verification model from AIG and SSA through 
in-tandem symbolic execution of HW and FW components.

\textbf{Step 2: Bounded Model Checking}
%
The co-verification model is checked with bounded model checking (BMC).
Given an unwind depth depth~$k$ and a co-design property $P$, BMC operates
by unwinding the co-verification model~$M$ up to depth~$k$ starting from
initial state $x_0$, represented by an initial state predicate~$I$.
This results in the following formula which is then checked for
satisfiability using an efficient SAT or SMT procedure.
%
\[ I(x_0) \wedge \bigwedge_{i=0}^{k-1} M(x_i, x_{i+1}) \wedge \bigvee_{i=0}^{k} 
\neg P(x_i) \]
%
Thus, BMC exploits the finiteness of $M$ and creates k copies of the
$M$ for each unrolling.

%\hwcbmcv provides specific handshake primitives such
%as $next\_timeframe()$ and $set\_inputs()$ to model FW-HW communications.

\textbf{Step 3: Scenario/Transaction Pair} In \hwcbmcv, the symbolic execution 
of HW and FW components meet only at the solver phase.  Thus, a complex
monolithic formula is passed on to the solver for verification purpose.  This
limits the opportunity for extracting relevant scenario/transaction pairs
through symbolic execution. Instead, the technique relies on the underlying
SAT/SMT solver to effeciently reason about the co-verification model. 
%
\Omit{
The hardware and firmware run independently of each other. 
The communication between them takes place through these function calls. 
}
%===============================================================================
\section{Property Specification for Co-verification}\label{property}
%===============================================================================
Lack of support for property specification in a HW/SW co-design is one of the
stumbling blocks for the application of formal techniques in co-verification.  
In this paper, we express a co-design property in C 
with the support of special primitive, called {\em next\_timeframe()}, 
that can express temporal correlation between HW and FW events. 
In this paper, we classify co-design properties into two 
classes -- {\em transaction-level} property and {\em component-level} property.
%-------------------------------------------------------------------------------
\subsection{Transaction-level Property}
%-------------------------------------------------------------------------------
A {\em Transaction-level} property ($\psi_{t}$) formally captures the 
design intent of the interaction between a $\mathcal{ST}$ and $\mathcal{HT}$, 
that is, it is defined at the transaction boundaries.  
Recall that a $\mathcal{ST}$ is a service in response to some external environment 
or HW interrupts (for example acknowledgement or exceptions). Similarly, 
$\mathcal{HT}$ is a finite-length service in response to some SW events or 
other HW events.

>>>>

\end{enumerate}
%
%===============================================================================
\section{Working Example}
%{\bf Running example: }
%===============================================================================
In this section, we present a working example of FW and HW model to explain our 
co-verification methodologies.

@@

\textbf{Co-design Property} We verify that the transmitted data is the same as the 
received data in the {\em loopback} mode.  To do so, we annotate the FW model 
with an assertion $assert()$, as shown in Figure~\ref{firmware}.  Note that the 
array containers $rx\_b$ and $tx\_b$ stores the data obtained through the $inb$
and $outb$ function calls which directly read from and write to the UART IP ports
respectively. Section~\ref{property} describes the property specification in
HW-FW co-designs in more detail. 
\[
  for(i=0; i<=N; i++) 
   \quad assert(rx\_b[i] == tx\_b[i]);
\]
%
\section{Scenario and Transaction}
%
We describe the notion of \emph{transaction} and \emph{scenario} with the
help of our working example. \\
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{|l|l|}
\hline
Wishbone Interface & Main Module \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
typedef unsigned char u8;
unsigned char inb 
 (unsigned long port) {
  return wb_read(port);
}
void outb (u8 value, 
 unsigned long port) {
  wb_write(port, value);
}
void wb_reset(void) {
 rst_i = 1;
 set_inputs();
 next_timeframe();
 rst_i = 0;
 stb_i = 0; 
 cyc_i = 0;
}
void wb_idle() {
 set_inputs();
 next_timeframe();
}
void wb_write(_u32 addr, 
     _u8 b) {
 adr_i = addr;
 dat_i = b;
 we_i = 1;
 cyc_i = 1;
 stb_i = 1;
 set_inputs();
 next_timeframe();
 we_i = 0;
 cyc_i = 0;
 stb_i = 0;
}
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]
int main() {
wb_reset();
wb_idle();
// Configure the uart
outb (0x13, UART_MC);  
outb (0x80, UART_CM3);
outb (0x00, UART_CM2);
outb (0x00, UART_CM1);
outb (0x00, UART_CR);  
outb (0x03, UART_IE); 
char tx_b[] = "Hello world";
_u8 status = 0;
char rx_b[100];
int i=0,c=0,d=0;
// data transfer in loopback
for (i=0; i<1990; i++){
if (irq_o){
 status=inb(UART_IS)&0x0c;
 if(istatus==0x0c){
 //it was a tx_empty interrupt
 outb(*(tx_b+c),UART_TR); c++;
 }
 else{ //status==0x04
 //it was an rx_data interrupt
  rx_b[d] = inb(UART_TR); d++;
 }
}else {
 // no interrupt. 
 wb_idle();
 wb_idle();
 }
}
// property
for(i=0; i<=10; i++)
 assert(rx_b[i] == tx_b[i]);
}
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{Firmware model of UART IP}
\label{firmware} 
\end{figure}
%

\textbf{Scenario ($\mathcal{S}$):}
Let $B$ denote the set of all possible behaviors of a system. A scenario is a set of 
constraints $\mathcal{S}$ that restrict the system behaviors
to $B' \subseteq B$. For example, the left hand side of Figure~\ref{fig:st} 
gives an example scenario for the UART FW shown in Figure~\ref{firmware}. 
Here, $\mathcal{S}$ configures the UART IP in {\em loopback} mode with 
interrupt enabled.

@@

\textbf{Firmware Transaction ($\mathcal{FT}$):} 
A Firmware Transaction ($\mathcal{FT}$) is an untimed state machine 
with a start state and an end state, which is similar to the notion of 
a transaction in TLM~\cite{codes14}. 
Each transition in $\mathcal{FT}$ is guarded with a quantifier free predicate
over FW variables and HW signals.
If the guard evaluates to $TRUE$, it may perform some action, which may  
read/write data. If the guard is not specified, it is assumed to be $TRUE$.
$\mathcal{FT}$ is often manifested as control-flow paths in the firmware. 
The right hand side of Figure~\ref{fig:st} gives a 
$\mathcal{FT}$ for the UART firmware code shown in Figure~\ref{firmware}. 
It has a start state and an end state. $\mathcal{FT}$ is triggered from 
UART output signal, $irq_o$. An execution of $\mathcal{FT}$, also called a 
FW transaction instance, takes a simple path from the start state to the end state. 
The path taken by an execution is determined by the state of 
the FW variables or its interacting HW signals. Repeated execution of 
$\mathcal{FT}$ can happen only in a sequential manner \sjcmt{not clear what this means}.

@@

Let $\mathcal{G}$ be the set of quantifier free boolean predicates over 
state elements, $\{Reg\}$, $\{Mem\}$, $\{Wire\}$.  
A {\em transaction step}, $\varphi$, is defined by a 2-tuple,
$\varphi \equiv \langle \Gamma_{\varphi}, \Delta_{\varphi} \rangle$, where 
$\Gamma_{\varphi} \equiv \{\alpha_1, \alpha_2, \ldots \alpha_{m}\}$
is a set of concurrent state updates to $\{Reg\}$, $\{Mem\}$, $\{Wire\}$ 
and $\Delta_{\varphi} \equiv \{\beta_1, \beta_2, \ldots \beta_{n}\} \subseteq
\mathcal{G}$.

A {\em Hardware Transaction} $(\mathcal{HT}) = (V, \Sigma, \rightarrow, init,
comp)$ is a labelled transition system, where $V$ is a set of transaction steps
$\varphi$ with a unique source vertex, $init \in V$, and a unique sink vertex, 
$comp \in V$, $\Sigma$ is a set of labels such that $\Sigma \subseteq 2^{G}$ \sjcmt{2 power G may not be required, just G may be sufficient as it is closed under disjunction, conjunction and negation} and  
$\rightarrow \subseteq V \times \Sigma \times V$ is a set of labelled
transitions.

@@

\caption{Proposed HW/SW Co-verification Flow

@@

\textbf{Step 3: Co-verification Model}
%
An unified co-verification model is constructed through appropriate composition of 
FW and its interacting HW threads.  This is done by instrumenting the 
software netlist model of HW RTL inside the FW thread at every invocation 
of the $next\_timeframe()$ primitive in interface functions.  Recall that a
call to $next\_timeframe()$ is used to advance the HW clock by one cycle.  A call
to the software netlist emulates the effect of one clock period in the RTL 
model. That is, the state of state holding elements (registers) and combinational 
elements (wires) are uniform in software netlist and RTL model after every invocation of 
$next_timeframe()$.  Furthermore, a software
netlist is a sequential model of HW RTL which is represented in C.  Thus, the
composition leads to a single-threaded sequential co-verification model in C.  
%
We now formally define a our co-verification model, denoted by 
$\mathcal{M}_{seq}$.
%
\[ 
\begin{array}[t]{@{}lll}
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\
\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{moduleName(Var_1,\dots,Var_n)} \\
\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\ 
\end{array}
\]
%
Consider sets of expressions $Exp$ and Boolean expressions $BExp$ 
over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 
$(b \in BExp)$. A $\mathit{CStatement}$ is an assignment,
conditional, sequential concatenation or a loop.
$\mathit{moduleName(Var_1,\dots,Var_n)}$ invokes the top level 
procedure of $\mathcal{SN}$, which is also the top level 
module of the HW RTL, since the translation from RTL to software preserves
the module hierarchy.  The interface variables $\{Var\_1,\dots,Var\_n\}$ 
(input, output and inout ports) are declared inside the 
FW to communicate with $\mathcal{SN}$.  Figure~\ref{compose} shows the 
composition of $\mathcal{SN}$ and FW thread for the UART IP.
%
%\begin{figure}[t]
%\scalebox{.50}{\import{figures/}{compose.pspdftex}}
%  \caption{Composition of FW and HW threads}
%\label{fig:interleaving2}
%\end{figure}
%

\textbf{Step 4: Property Driven Slicing of Co-verification Model}
A property-driven slicing is performed on the unified model, $\mathcal{M}_{seq}$.  
This step is purely syntactic, meaning that we perform a backward dependency 
analysis starting from the property which only preserve those 
program fragments that are relevant to a given property. The sliced program 
is then passed to the symbolic execution engine for path-based exploration. \\
%
%=======================================================================================
%\subsection{\verifox: Path-wise Symbolic Execution Tool} 
%=======================================================================================
\input{verifox}
%
\subsection{Approach 2: Monolithic Co-verification Technique}
%
\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.6]{figures/traditional_flow.pdf}
\caption{Monolithic HW/SW Co-verification Flow}
\label{fig:conventional}
\end{center}
\end{figure}
%
Figure~\ref{fig:conventional} illustrates the monolithic approach to HW/SW
co-veri\-fi\-cation, as implemented in \hwcbmcv.  
%\hwcbmcv is the only open-source
%formal co-verification tool that supports Verilog RTL (System Verilog 2009
%standard) and C/C++ firmware.  
In contrast to path-based approach,
\hwcbmcv maintains two separate flows for HW and FW.  The top flow in
Figure~\ref{fig:conventional} uses synthesis to obtain either a bit-level or
a word-level netlist from Verilog RTL.  A bit-level netlist is represented 
in And Inverter Graph (AIG) format.  Whereas, a word-level netlist is represented 
in format that resembles SMT-LIB standards.  The bottom flow illustrates the
translation of C program into static single assignment (SSA) form. 
Synthesizing netlist from RTL is a standard technique which is widely 
used in hardware model checkers~\cite{}. Similarly, translating a C program 
into a SSA representation is a standard technique which is widely used in 
software verification tools, such as CBMC~\cite{}, 2LS~{}. 
These two flows meet only at the solver.  Thus, \hwcbmcv generates a
unified monolithic co-verification model from AIG and SSA which is 
represented in Conjunctive Normal Form (CNF).  This model is then used to perform 
bounded model checking~\cite{} using the underlying SAT/SMT solvers.  
\hwcbmcv provides specific handshake primitives such
as $next\_timeframe()$ and $set\_inputs()$ to model FW-HW communications.

\rmcmt{
To this end, \hwcbmcv constructs a co-verification model through 
in tandem symbolic execution of the HW and FW models.  A bit-level 
netlist representation in And Inverter Graph (AIG) format or word-level netlist in
format similar to SMT-LIB standards is synthesized from a HW Register Transfer 
Level (RTL) implementation. Similarly, the SW, given in C, is automatically 
translated into a Static Single Assignment (SSA) form.  A monolithic verification 
condition is generated from these two models which is then solved using a 
}
%
\Omit{
The hardware and firmware run independently of each other. 
The communication between them takes place through these function calls. 
}
%===============================================================================
\section{Property Specification for Co-verification}\label{property}
%===============================================================================
Lack of support for property specification in a HW/SW co-design is one of the
stumbling blocks for the application of formal techniques in co-verification.  
In this paper, we express a co-design property in C 
with the support of special primitive, called {\em next\_timeframe()}, 
that can express temporal correlation between HW and FW events. 
In this paper, we classify co-design properties into two 
classes -- {\em transaction-level} property and {\em component-level} property.
%-------------------------------------------------------------------------------
\subsection{Transaction-level Property}
%-------------------------------------------------------------------------------
A {\em Transaction-level} property ($\psi_{t}$) formally captures the 
design intent of the interaction between a $\mathcal{ST}$ and $\mathcal{HT}$, 
that is, it is defined at the transaction boundaries.  
Recall that a $\mathcal{ST}$ is a service in response to some external environment 
or HW interrupts (for example acknowledgement or exceptions). Similarly, 
$\mathcal{HT}$ is a finite-length service in response to some SW events or 
other HW events.
